\documentclass[11pt,article,oneside]{memoir}
\usepackage{org-preamble-xelatex}
\input{vc}



\title{\bigskip \bigskip Acknowledgments}
 
\author{\bigskip\Large }

\begin{document}  

\setsansfont[Mapping=tex-text, BoldFont={* Bold SemiCondensed}, ItalicFont={* Semibold SemiCondensed Italic}]{Myriad Pro}
\setmonofont[Mapping=tex-text,Scale=MatchLowercase]{Consolas}
\setromanfont[Mapping=tex-text,Numbers=OldStyle]{Minion Pro}


\setkeys{Gin}{width=1\textwidth}  
\setromanfont[Mapping=tex-text,Numbers=OldStyle]{Minion Pro} 
\setsansfont[Mapping=tex-text]{Minion Pro} 
\setmonofont[Mapping=tex-text,Scale=0.8]{Consolas}
\chapterstyle{article-4} 
\pagestyle{kjh}

\published{Draft only. Please do not cite without permission.}

\maketitle


My graduate school colleagues at the University of Nebraska-Lincoln were
invaluable resources for knowledge, questions, and camaraderie. Michelle
Tiedje, Brandon Locke, Brent Rogers, Brian Sarnacki, Robert Jordan, Andy
Wilson, Rebecca Wingo, \ldots{}.. Special thanks also to Laura Weakly,
Karin Dalziel, Keith Nickum, and Catherine Walter.

At Stanford, I have benefited from another rich community of ideas and
thinkers. My colleagues have been wonderful friends, inquisitive
thinkers, and indefinitely helpful. Maria \_\_, Glauco \emph{\_}, Nicole
Coleman, Mikael Wolfe, Katie McDonough, Cameron Blevins, Richard White,
Zephyr Frank, Nick Bauch, have all served as sounding boards for my
ideas. I'm grateful for the time you've offered to listen to my ideas,
the probing questions you've asked, and the encouragement you've
provided.


\end{document}
